Automatic Visual Inference for Linear Regression

### 2022-03-15 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
finish cubic

# Simulation Setup

<!-- This project focuses on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model. To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots. -->
<!-- Because the first project is still ongoing, we consider the result we have is not enough to write as a paper at the moment. Hence, the materials provided in this chapter are two main parts of the draft paper that will contain enough details for understanding the project. -->

#### To train computer vision models for evaluating lineups, we need a large amount of data.

#### Meanwhile, we need to set up a comparison between the computer vision models and humans on the ability of reading residual plot for understanding the performance of the models.

>>>>

# Simulation Setup - Cubic Model

#### To train computer vision models, we use simulation data generated from two linear regression model with **non-linearity** and **heteroskedasticity** issues.

#### The first model is a **cubic model** which can be expressed by: $$\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},$$ where $c \in (0,2)$, $a \in (-3,3)$, $b \in (-3,3)$, $\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})$, $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$ are $n\times1$ matrices.

#### In the simulation, $X_i$, $i=1,...,n$, has equal chance to follow one of the following distributions: $N(0, 0.09)$, $U(-1, 1)$, $Lognormal(0,0.36)/3 - 1$ and $-Lognormal(0,0.36)/3 + 1$. These distributions are chosen such that most the realizations will fall between $-1$ and $1$.

#### The distribution of $Z_i$, $i=1,...,n$, has $50$% chance to be a uniform distribution ranged from $-1$ to $1$, and $50$% chance to be a discrete uniform distribution with $z_n$ outcomes simulated from a uniform distribution ranged from $-1$ to $1$. $z_n$ itself is a discrete uniform distribution with outcomes $\{x: 9<x<21 \text{ and } x \in \mathbb{Z} \}$, which defines the number of possible values $Z_i$ could take.

# Simulation Setup - Cubic Model

#### The simulated data is fitted by OLS with equation: $$\boldsymbol{Y}= \beta_0 + \beta_1\boldsymbol{X} + \beta_2\boldsymbol{Z} +\boldsymbol{u},$$ where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})$.

#### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form. Then, the **residuals** obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.

#### The **effect size** is derived as: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

.pull-left[
```{r}
mod <- cubic_model(a = 5, b = 10, c = 1)
mod$plot_lineup(mod$gen_lineup(300), remove_grid_line = TRUE, theme = theme_light()) + xlab("Fitted values") + ylab("Residuals")
```

.caption[
Fig. 8: _A lineup of residual plots generated with the cubic model._ 
]
]

.pull-right[
```{r}
mod <- cubic_model(-1, -1, 1, 0.5)
dat <- mod$gen(500, fit_model = TRUE)
mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p1

mod <- cubic_model(-2, -1, 1, 0.5)
dat <- mod$gen(500, fit_model = TRUE)
mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p2

mod <- cubic_model(-5, -5, 1, 0.5)
dat <- mod$gen(500, fit_model = TRUE)
mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p3

mod <- cubic_model(-20, -20, 1, 0.5)
dat <- mod$gen(500, fit_model = TRUE)
mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p4

list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=2)
```

.caption[
Fig. 9: _Residual plots of cubic model under different effect sizes._ 
]
]

# Simulation Setup - Heteroskedasticity Model

#### The second model is a heteroskedasticy model which can be expressed by: $$Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,$$ where $a \in \{-1,0,1\}$, $b\in (0,32)$ and $\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)$.

#### $\boldsymbol{X}$ could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted log-normal distribution, adjusted negative log-normal distribution and discrete uniform distribution.

# Simulation Setup - Heteroskedasticity Model

#### The simulated data is fitted by OLS with equation: $$\boldsymbol{Y}= \beta_0 + \beta_1\boldsymbol{X} +\boldsymbol{u},$$ where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})$.

#### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form. The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
